A fast unified model for parsing and sentence understanding

### 2016-03-16 writing/hard_stack_paper/paper.tex
Results tweaks.
Writing up CP's comments.

Of these, it is easy to argue that the TreeRNN is the principled choice, since meaning in natural language sentences constructed incrementally according to a tree structure \citep{Partee84,Janssen97}. However, while these models have shown promise \citep{tai2015improved,li2015tree,bowman2015trees}, they have largely been overlooked in favor of sequence-based RNNs because of their incompatibility with batched computation and their reliance on external parsers.  Batched computation---performing synchronized computation across many examples at once---yields order-of-magnitude improvements in model run time, and is a crucial enabling technology in allowing neural networks to be trained efficiently on large datasets. Because TreeRNNs use a different model structure for each sentence, batched computation is impossible in standard implementations. In addition, standard TreeRNN models can only operate on sentences that have already been processed by a syntactic parser, which slows and complicates the implementation of these models for most applications.

This paper introduces a new model to address both these issues: the Stack-augmented Parser-Interpreter Neural Network, or SPINN, shown in Figure~\ref{fig:m1-views}. The SPINN executes the computations of a tree-structured model in a linearized sequence of operations, and can incorporate a neural network parser that determines the overall tree structure on-the-fly. This design improves upon the TreeRNN architecture in three ways:
\begin{itemize}
\item At test time, it can simultaneously parse and interpret unparsed sentences, removing the dependence on an external parser.
\item It supports batched computation for both parsed and unparsed sentences, supporting dramatic speedups over standard TreeRNNs.
\item It supports a novel tree-sequence hybrid mechanism for handling local context in sentence interpretation that yields substantial gains in accuracy over either pure sequence-based models or pure tree-based models.
\end{itemize}

We evaluate the SPINN on the Stanford Natural Language Inference sentence understanding task \citep[SNLI,][]{snli:emnlp2015}, and find that it significantly outperforms other sentence encoding-based models, and that it yields speed increases of upwards of 10x over comparable TreeRNN models.

\subsection{Related work}

There has been a fairly long history of work on building neural-network based parsers that use the core operations and data structures from transition-based parsing, of which shift-reduce parsing is a version \citep{henderson2004discriminative,emami2005neural,titov2010latent,chen2014,buys2generative,dyer-EtAl:2015:ACL-IJCNLP,kiperwasser2016easy}. In addition, there has been recent work \citep{zhang2016top,dyer2016rnn} proposing models designed primarily for generative language modeling tasks that use these structures as well. To our knowledge, the SPINN is the first model to use these structures for the purpose of sentence interpretation, rather than parsing or generation.

@@

\section{Our model: The SPINN}

@@

The SPINN is inspired by the shift-reduce parsing formalism \citep{aho1972theory}, which builds a tree structure over a sequence (e.g. a natural language sentence) by a single left-to-right scan over the tokens of the sequence. Shift-reduce parsing is frequently used in natural language parsing \citep[e.g.][]{nivre2003efficient}.

@@

The core SPINN model implements a sequential shift-reduce parser which operates over sentences. It is designed to produce a vector representation of the sentence as output, as opposed to a tree as in standard shift-reduce parsing. This relies on two key modifications to the shift-reduce algorithm:
\begin{itemize}
\item The parser's intermediate representations, including the representations of partial tree structures on the stack, are fixed-length vectors.
\item The \reduce~operation combines two such representations of trees on the stack into a larger tree using a version of the TreeLSTM neural network layer \citep{tai2015improved}.
\end{itemize}

This subsection describes the core functions and representations of the SPINN. We begin with the most basic model, SPINN-NP (for \textit{no parser}), which processes input sentences using parses provided by some external parser.

@@

The buffer is populated with the words in each sentence before the first step of computation. We use word representations from the standard 300-dimensional vector package provided with GloVe \citep{pennington2014glove}. We do not update these representations during training. Instead, we use a learned linear transformation to transform the representation of each input word $\vec{x}$ into a pair of vectors $\langle \vec{h}, \vec{c}\rangle$ that are stored side-by-side in the buffer and used as inputs to the TreeLSTM composition function:

\begin{equation}
\colvec{2}
    {\vec{h}}
    {\vec{c}}
= W_{\text{wd}} \vec{x}
\end{equation}

\paragraph{The composition function}
When a \reduce~operation is performed, vector representations of two tree nodes are popped off of the head of the stack and fed into a {\ii composition function}, which is a neural network function that produces a representation for a new tree node that is the parent of the two popped nodes. This new node is then pushed on to the stack.

Our composition is a version of the TreeLSTM of \citet{tai2015improved}. The TreeLSTM layer, which generalizes the LSTM neural network layer to tree- rather than sequence-based inputs, and it shares with the LSTM the idea of representing intermediate states as a pair of a fast-changing state representation $\vec{h}$ and a slower-changing memory representation $\vec{c}$. It is formulated as:

\begin{gather}
\colvec{4}
    {\vec{i}}
    {\vec{f}_l}
    {\vec{f}_r}
    {\vec{o}}
= \sigma\left(
W_{\text{iffo}}
\colvec{3}
    {\vec{h}_s^1}
    {\vec{h}_s^2}
    {\vec{e}}
\right) \label{eqn:lstm1}
\\
\vec{g}
= \tanh\left(
W_{g}
\colvec{3}
    {\vec{h}_s^1}
    {\vec{h}_s^2}
    {\vec{e}}
\right) \label{eqn:lstm2}
\\
\vec{c} = \vec{f}_l \odot \vec{c}_s^{\,2} + \vec{f}_r \odot \vec{c}_s^{\,1} + \vec{i} \odot \vec{g}  
\\
\vec{h} = \vec{o} \odot \vec{c}
\end{gather}

The results of this function are the pair $\langle\vec{h}, \vec{c}\rangle$, which are placed back on the stack. The two input tree nodes popped off the stack are represented as the pairs $\langle\vec{h}^1_s, \vec{c}^{\,1}_s\rangle$ and $\langle\vec{h}^2_s, \vec{c}^{\,2}_s\rangle$. In addition, $\vec{e}$ is an optional input argument which is either the empty vector or a vector from an external source like the tracking LSTM (see Section~\ref{sec:tracking}). $\odot$ denotes the elementwise product. Each vector-valued variable listed here is of the same dimension $D$ except $\vec{e}$, which is of the independent dimension $D_t$.

@@

This paper evaluates the SPINN on a sentence pair classification task. To classify a sentence pair, a feature vector $\vec{x}_{\text{classifier}}$ is first constructed. This feature vector is based on the final representations of each of the two sentences---the representations that appear at the head of the stack for each sentence after the final transition. In particular, the $\vec{h}$ portions of these representations are used. The final feature vector consists of the concatenation of these two vectors, $\vec{h}_{\text{premise}}$ and $\vec{h}_{\text{hypothesis}}$, their difference, and their elementwise product:

\begin{equation}
\vec{x}_{\text{classifier}} = 
\colvec{4}
    {\vec{h}_{\text{premise}}}
    {\vec{h}_{\text{hypothesis}}}
    {\vec{h}_{\text{premise}} - \vec{h}_{\text{hypothesis}}}
    {\vec{h}_{\text{premise}} \odot \vec{h}_{\text{hypothesis}}}
\end{equation}

Following \citep{snli:emnlp2015}, this feature vector is then passed to a series of ReLU neural network layers (the number of layers is a tuned hyperparameter), then passed into a linear transformation, and then finally passed to a softmax layer, which yields a distribution among the three labels. The model is trained in part using a cross-entropy objective over this distribution.

@@

The tracking LSTM takes three vectors as inputs at each step (highlighted in yellow in Figure~\ref{fig:m1-views}): the top element of the buffer $\vec{h}_b^1$, which would be moved in a \shift~operation, and the top two elements of the stack $\vec{h}_s^1$ and $\vec{h}_s^2$ which would be composed in a \reduce~operation.

\paragraph{Tracking left context for parsing} The decision function which determines which transitions (and thereby, which parses) the SPINN will use needs to take as its input some representation of the current state of the model. We use the hidden state of the tracking LSTM as that input. The tracking LSTM's limited access to the buffer makes this fast to compute, but somewhat more limited than the rich feature sets used in state-of-the-art parsers.

@@

It would be straightforward to augment the SPINN to support the use of some amount of right context as well, but this would add complexity to the model that we think is largely unnecessary: humans are very effective at understanding the beginnings of sentences before having seen or heard the ends, suggesting that it is possible to get by without the unavailable right context.

@@

The model described so far is the SPINN-NP model, which uses sequences of \shift~and \reduce~operations from an external parser. To build the fully independent SPINN model, we simply add a single-layer neural network classifier that chooses which operation to perform at each step using the state of the tracking LSTM:
\begin{equation}
\vec{p}_{\text{trans}} = \text{softmax}(W_{\text{trans}}\vec{h}_{\text{track}})
\end{equation}

The resulting vector $\vec{p}_{\text{trans}}$ is a probability distribution over parsing decisions (i.e., predicting whether to \shift~or \reduce~at the current timestep). In the full SPINN model, we follow whichever transition has a higher predicted probability in this distribution. This prediction module is trained to mimic the decisions of an external parser. In particular, we use the transition sequences corresponding to the constituency parses included in the SNLI corpus. The classifier is trained using a cross-entropy objective:
\begin{equation}
  L_{\text{trans}} = - \log \vec{p}_{\text{trans}}[t^*]
\end{equation}
where $t^*$ is the correct transition specified by the gold parse, and $\vec{p}[\cdot]$ represents vector indexing.

@@

\subsection{Implementing the SPINN}

\paragraph{The full objective function} Our objective function combines a cross-entropy objective $\mathcal{L}_{\text{s}}$ for the SNLI classification task, a cross-entropy objective $\{\mathcal{L}_p^t, \mathcal{L}_h^t\}$ for each parsing decision for each of the sentence and the hypothesis sentence at each step $t$, and an L2 regularization term on the trained parameters. The terms are weighted using the tuned hyperparameters $\alpha$ and $\lambda$:

\begin{equation}
\begin{split}
\mathcal{L}_{\text{m}} = &\mathcal{L}_{\text{s}} + \alpha \sum_{t=0}^{T-1} (\mathcal{L}_{\text{p}}^{t} + \mathcal{L}_{\text{h}}^{t}) + \lambda \|\theta\|^2_2
\end{split}
\end{equation}

\paragraph{Preparing the data} At training time, the SPINN requires both a transition sequences $\mathbf t$  and a token sequence $\mathbf x$ as its inputs for each sentence. The token sequence is simply the words in the sentence in order. $\mathbf t$ can be obtained from any constituency parse for the sentence by first converting that parse into an unlabeled binary parse, then linearizing it (with the usual in-order traversal), then taking each word token as a \shift~transition and each `)' as a \reduce~transition, as here:

@@

A naive implementation of the SPINN model would require simulating a stack of size $N$ for each input sentence. In order to update the model weights via backpropagation, we would also need to maintain all intermediate stack representations for later chain rule calculations. This implies a per-example space requirement of $N \times T \times D$, which is prohibitively large for significant batch sizes $M$ or sentence lengths $N$. (Recall that $T = 2N - 1$ is the number of transitions in the shift-reduce parsing sequence, and $D$ is the dimension of the vectors on the stack.) Such a naive implementation would also require duplicating a largely unchanged stack at each timestep (as \shift~and \reduce~operations only affect the top of the stack).

@@

In its bare form, before the addition of the tracking LSTM, the SPINN (or, SPINN-NP-NT, for \textit{no parsing, no tracking}) is precisely equivalent to a conventional tree-structured neural network model in the function that it computes, and it therefore also has the same learning dynamics. In both, the representation of each sentence consists of the representation of the words, combined recursively using a TreeRNN composition function (in our case, the TreeLSTM function).

@@

\begin{table*}[t]
  \centering\small
  \begin{tabular}{lcccc} 
    \toprule
Model                   & Params.    & Trans. acc.  &   Train  &   Test \\
\midrule
\multicolumn{5}{c}{Previous non-NN results}\\
\midrule
Lexicalized classifier \citep{snli:emnlp2015}
                        & --                & --                    &   99.7   &   78.2      \\
\midrule
\multicolumn{5}{c}{Previous sentence encoder-based NN results}\\
\midrule
100D LSTM encoders \citep{snli:emnlp2015}
                        & 221k               & --               &   84.8   &   77.6      \\
1024D pretrained GRU encoders \citep{DBLP:journals/corr/VendrovKFU15}
                        & 15m                & --              &   98.8   &   81.4       \\
300D Tree-based CNN encoders \citep{mou2015recognizing}
                        & 3.5m                & --             &   83.4   &   82.1       \\
\midrule
\multicolumn{5}{c}{Our results}\\
\midrule
300D LSTM RNN encoders          & 3.0m                  & --                &   83.9      &   80.6       \\
300D SPINN-NP-NT encoders
                        & 3.4m                  & --                &   84.4      &   80.9       \\
300D SPINN-NP encoders
                        & 3.7m                  & --                &   89.2      &   \underline{83.2}       \\
\result{300D SPINN encoders }
                        & ??                  & 92.2+dv            &   ?    &   83.?      \\          
% \result{300D SPINN-NP, sequence-based attn.   }     
%                         & ?                  & --                &   ?      &   86.3dv         \\
% \result{300D SPINN-NP, tree-based attn. }           
%                         & ?                  & --                &   ?      &   \textbf{?}\\
\midrule
\multicolumn{5}{c}{Other previous NN results}\\
\midrule
100D LSTM w/ word-by-word attention \citep{rocktaschel2015reasoning}
                        & 252k               & --              &   85.3   &   83.5       \\
300D mLSTM word-by-word attention model \citep{DBLP:journals/corr/WangJ15b}
                        & 1.9m               & --             &   92.0   &   86.1      \\
300D LSTMN with deep attention fusion \citep{cheng2016long}
                        & 1.4m               & --                &   92.3   &   \textbf{89.1}      \\
    \bottomrule
  \end{tabular}
\caption{\protect\label{tab:results}Results on SNLI 3-way inference classification. Params. is the approximate number of trained parameters (excluding word embeddings for models where they are trained). Trans. acc. is the model's accuracy in predicting parsing transitions. Train and test are SNLI classification accuracy.} 
\end{table*}

@@

We evaluate the SPINN on the task of natural language inference \citep[NLI, also known as recognizing textual entailment, or RTE][]{dagan2006pascal,MacCartney09}. NLI is a sentence pair classification task, in which a model reads two sentences (a premise and a hypothesis), and outputs a judgment of {\it entailment}, {\it contradiction}, or {\it neutral}, reflecting the relationship between the meanings of the two sentences, as in this example from the Stanford NLI corpus \citep[SNLI][]{snli:emnlp2015}, which we use for training and evaluation:

@@

Even though NLI is framed as a relatively simple three-way classification task, it is nonetheless an effective way of evaluating the ability of some model to extract broadly informative representations of sentence meaning. In order for a model to perform reliably well on NLI across a range of sentence types and text genres, it must be able to represent and reason with all of the core phenomena of natural language semantics, including quantification, coreference, scope, and several types of ambiguity.

@@

We evaluate four models on the SNLI task. Each uses 300D hidden states:
\begin{itemize}
\item A baseline LSTM model (similar to that of \citet{snli:emnlp2015}) that uses the same classifier architecture as our models, but encodes sentences using a single layer LSTM-RNN sequence model.
\item The minimal SPINN-NP-NT (\textit{no parsing, no tracking}) model, which uses transitions from an external parser and has no tracking LSTM to inform semantic composition. This model is equivalent to a TreeLSTM \citep{tai2015improved}.
\item The SPINN-NP model, which adds the tracking LSTM to guide composition, making it a hybrid tree-sequence model.
\item The full SPINN model, which makes its own parsing decisions, and does not depend on an external parser at test time.
\end{itemize}

We compare our models against a range of baselines, including the strongest published non-neural network-based result from \citep{snli:emnlp2015}, a set of previous neural network models built around sentence encoders, and a set of models built around soft attention architectures. We should note that these attention-based models are by far the strongest published on SNLI to date, but these are not necessarily comparable to ours: SNLI was originally presented as an evaluation and development task for sentence encoding models, on the premise that sentence encoding models that do well on SNLI capture sentence meaning in a general sense, and are well suited to use on a range of NLP tasks. The attentional models are specifically adapted to the sentence-pair structure of SNLI examples and cannot act as such general sentence encoders. We focus on models which do support general sentence encoding in our evaluation.

@@

We find that the bare SPINN-NP-NT model performs little better than the RNN baseline, but that the SPINN-NP with the added tracking LSTM reaches state-of-the-art accuracy among sentence encoding based models. The success of the SPINN-NP model, which is a hybrid tree-sequence model, suggests that the tree- and sequence-based encoding methods are complementary. The full SPINN model with its relatively weak internal parser performs slightly less well, but nonetheless exceeds the performance of the LSTM baseline. None of our models reached the performance of the strongest attention-based models, which were explicitly engineered for the sentence pair-based structure of the SNLI task.

Both the SPINN-NP and the full SPINN outperform all previous sentence encoding models by a significant margin. Most notably our modes outperform the TBCNN of \citet{mou2015recognizing}, which also uses tree-structured composition for local feature extraction, but uses simpler pooling techniques to build sentence features in the interest of efficiency. Our results show that a model that uses tree-structured composition fully (SPINN) outperforms one which uses it only partially (TBCNN), which in turn outperforms one which does not use it at all (LSTM RNN).

@@

We implemented the feedforward of the SPINN-NP-NT model (which is equivalent to a standard TreeRNN) in C++/CUDA in order to compare to the performance of a standard CPU-based TreeRNN solution from \citet{irsoy2014deep}. Such a standard TreeRNN implementation operates on a single example at a time and is thus not feasible to run on a GPU.\footnote{We chose to reimplement and evaluate only the thin-stack feedforward (not training/backpropagation) in C++/CUDA, as this is the more critical performance metric for real-world applications.}

Figure~\ref{fig:speed} compares the runtime performance of the two models computing representations on data from the Stanford Sentiment Treebank \todo{cite}. Each model is restricted to run on sentences of 30 tokens or fewer (for the SPINN model, this means we iterate over a sequence of 59 transitions). We fix the model dimensionality and word embedding dimensionality at 300 for all tests.

@@

\todo{[JG] Note that the full SPINN (i.e. Model 1) is in the same complexity class as SPINN-NP (what we're evaluating here) with very similar constants.}

>>>>

Of these, it is easy to argue that the TreeRNN is the principled choice, since meaning in natural language sentences constructed incrementally according to a tree structure \citep{Partee84,Janssen97}. However, while these models have shown promise \citep{tai2015improved,li2015tree,bowman2015trees}, they have largely been overlooked in favor of sequence-based RNNs because of their incompatibility with batched computation and their reliance on external parsers.  Batched computation---performing synchronized computation across many examples at once---yields order-of-magnitude improvements in model run time, and is a crucial enabling technology in allowing neural networks to be trained efficiently on large datasets. Because TreeRNNs use a different model structure for each sentence, batched computation is impossible in standard implementations. In addition, standard TreeRNN models can only operate on sentences that have already been processed by a syntactic parser, which slows and complicates the use of these models at test time for most applications.

This paper introduces a new model to address both these issues: the Stack-augmented Parser-Interpreter Neural Network, or SPINN, shown in Figure~\ref{fig:m1-views}. SPINN executes the computations of a tree-structured model in a linearized sequence of operations, and can incorporate a neural network parser that determines the overall tree structure on-the-fly. This design improves upon the TreeRNN architecture in three ways:
\begin{itemize}
\item At test time, it can simultaneously parse and interpret unparsed sentences, removing the dependence on an external parser.
\item It supports batched computation for both parsed and unparsed sentences, supporting dramatic speedups over standard TreeRNNs.
\item It supports a novel tree-sequence hybrid mechanism for handling local context in sentence interpretation that yields substantial gains in accuracy over either pure sequence-based models or pure tree-based models.
\end{itemize}

We evaluate SPINN on the Stanford Natural Language Inference sentence understanding task \citep[SNLI,][]{snli:emnlp2015}, and find that it significantly outperforms other sentence encoding-based models, and that it yields speed increases of upwards of 10x over comparable TreeRNN models.

\section{Related work}

There has been a fairly long history of work on building neural-network based parsers that use the core operations and data structures from transition-based parsing, of which shift-reduce parsing is a version \citep{henderson2004discriminative,emami2005neural,titov2010latent,chen2014,buys2generative,dyer-EtAl:2015:ACL-IJCNLP,kiperwasser2016easy}. In addition, there has been recent work \citep{zhang2016top,dyer2016rnn} proposing models designed primarily for generative language modeling tasks that use these structures as well. To our knowledge, SPINN is the first model to use these structures for the purpose of sentence interpretation, rather than parsing or generation.

@@

\section{Our model: SPINN}

@@

SPINN is inspired by the shift-reduce parsing formalism \citep{aho1972theory}, which builds a tree structure over a sequence (e.g., a natural language sentence) by a single left-to-right scan over the tokens. Shift-reduce parsing is frequently used in natural language parsing \citep[e.g.,][]{nivre2003efficient}.

@@

The core SPINN model implements a sequential shift-reduce parser which operates over sentences. It is designed to produce a vector representation of the sentence as output, as opposed to a tree as in standard shift-reduce parsing. This relies on two key modifications to the shift-reduce algorithm:
\begin{itemize}
\item The parser's intermediate representations, including those of partial tree structures on the stack, are fixed-length vectors.
\item The \reduce~operation combines two such representations of trees on the stack into a larger tree using a version of the TreeLSTM neural network layer \citep{tai2015improved}.
\end{itemize}

This subsection describes the core functions and representations of SPINN. We begin with the most basic model, SPINN-NP (for \textit{no parser}), which processes input sentences using parses provided by some external parser.

@@

The buffer is populated with the words in each sentence before the first step of computation. We use word representations from the standard 300-dimensional vector package provided with GloVe \citep{pennington2014glove}. We do not update these representations during training. Instead, we use a learned linear transformation to transform the representation of each input word $\vec{x}$ into a pair of vectors $\langle \vec{h}, \vec{c}\rangle$ that are stored side-by-side in the buffer and used as inputs to the TreeLSTM composition function:
\begin{equation}
\colvec{2}
    {\vec{h}}
    {\vec{c}}
= W_{\text{wd}} \vec{x}
\end{equation}
\paragraph{The composition function}
When a \reduce~operation is performed, vector representations of two tree nodes are popped off of the head of the stack and fed into a {\it composition function}, which is a neural network function that produces a representation for a new tree node that is the parent of the two popped nodes. This new node is then pushed on to the stack.

Our composition is a version of the TreeLSTM of \citet{tai2015improved}. The TreeLSTM layer, which generalizes the LSTM neural network layer to tree- rather than sequence-based inputs, and it shares with the LSTM the idea of representing intermediate states as a pair of a fast-changing state representation $\vec{h}$ and a slower-changing memory representation $\vec{c}$. It is formulated as:
\begin{gather}
\colvec{4}
    {\vec{i}}
    {\vec{f}_l}
    {\vec{f}_r}
    {\vec{o}}
= \sigma\left(
W_{\text{iffo}}
\colvec{3}
    {\vec{h}_s^1}
    {\vec{h}_s^2}
    {\vec{e}}
\right) \label{eqn:lstm1}
\\
\vec{g}
= \tanh\left(
W_{g}
\colvec{3}
    {\vec{h}_s^1}
    {\vec{h}_s^2}
    {\vec{e}}
\right) \label{eqn:lstm2}
\\
\vec{c} = \vec{f}_l \odot \vec{c}_s^{\,2} + \vec{f}_r \odot \vec{c}_s^{\,1} + \vec{i} \odot \vec{g}  
\\
\vec{h} = \vec{o} \odot \vec{c}
\end{gather}
The results of this function are the pair $\langle\vec{h}, \vec{c}\rangle$, which are placed back on the stack. The two input tree nodes popped off the stack are represented as the pairs $\langle\vec{h}^1_s, \vec{c}^{\,1}_s\rangle$ and $\langle\vec{h}^2_s, \vec{c}^{\,2}_s\rangle$. In addition, $\vec{e}$ is an optional input argument which is either the empty vector or a vector from an external source like the tracking LSTM (see Section~\ref{sec:tracking}). $\odot$ denotes the elementwise product. Each vector-valued variable listed here is of the same dimension $D$ except $\vec{e}$, which is of the independent dimension $D_t$.

@@

This paper evaluates SPINN on a sentence pair classification task. To classify a sentence pair, a feature vector $\vec{x}_{\text{classifier}}$ is first constructed. This feature vector is based on the final representations of each of the two sentences---the representations that appear at the head of the stack for each sentence after the final transition. In particular, the $\vec{h}$ portions of these representations are used. The final feature vector consists of the concatenation of these two vectors, $\vec{h}_{\text{premise}}$ and $\vec{h}_{\text{hypothesis}}$, their difference, and their elementwise product:
\begin{equation}
\vec{x}_{\text{classifier}} = 
\colvec{4}
    {\vec{h}_{\text{premise}}}
    {\vec{h}_{\text{hypothesis}}}
    {\vec{h}_{\text{premise}} - \vec{h}_{\text{hypothesis}}}
    {\vec{h}_{\text{premise}} \odot \vec{h}_{\text{hypothesis}}}
\end{equation}
Following \citet{snli:emnlp2015}, this feature vector is then passed to a series of ReLU neural network layers (the number of layers is a tuned hyperparameter), then passed into a linear transformation, and then finally passed to a softmax layer, which yields a distribution among the three labels. The model is trained in part using a cross-entropy objective over this distribution.

@@

The tracking LSTM takes three vectors as inputs at each step (highlighted in yellow in Figure~\ref{fig:m1-views}): the top element of the buffer $\vec{h}_b^1$, which would be moved in a \shift~operation, and the top two elements of the stack $\vec{h}_s^1$ and $\vec{h}_s^2$, which would be composed in a \reduce~operation.

\paragraph{Tracking left context for parsing} The decision function which determines which transitions (and thereby, which parses) SPINN will use needs to take as its input some representation of the current state of the model. We use the hidden state of the tracking LSTM as that input. The tracking LSTM's limited access to the buffer makes this fast to compute, but somewhat more limited than the rich feature sets used in state-of-the-art parsers.

@@

It would be straightforward to augment SPINN to support the use of some amount of right context as well, but this would add complexity to the model that we think is largely unnecessary: humans are very effective at understanding the beginnings of sentences before having seen or heard the ends, suggesting that it is possible to get by without the unavailable right context.

@@

The model described so far is SPINN-NP (\ii{no parsing}), which uses sequences of \shift~and \reduce~operations from an external parser. To build the fully independent SPINN model, we simply add a single-layer neural network classifier that chooses which operation to perform at each step using the state of the tracking LSTM:
\begin{equation}
\vec{p}_{\text{trans}} = \text{softmax}(W_{\text{trans}}\vec{h}_{\text{track}})
\end{equation}
The resulting vector $\vec{p}_{\text{trans}}$ is a probability distribution over parsing decisions (i.e., predicting  \shift~or \reduce~at the current timestep). In the full SPINN model, we follow whichever transition has a higher predicted probability in this distribution. This prediction module is trained to mimic the decisions of an external parser. In particular, we use the transition sequences corresponding to the constituency parses included in the SNLI corpus. The classifier is trained using a cross-entropy objective:
\begin{equation}
  L_{\text{trans}} = - \log \vec{p}_{\text{trans}}[t^*]
\end{equation}
where $t^*$ is the correct transition specified by the gold parse, and $\vec{p}[\cdot]$ represents vector indexing.

@@

\subsection{Implementing SPINN}

\paragraph{The full objective function} Our objective function combines a cross-entropy objective $\mathcal{L}_{\text{s}}$ for the SNLI classification task, a cross-entropy objective $\{\mathcal{L}_p^t, \mathcal{L}_h^t\}$ for each parsing decision for each of the sentence and the hypothesis sentence at each step $t$, and an L2 regularization term on the trained parameters. The terms are weighted using the tuned hyperparameters $\alpha$ and $\lambda$:
\begin{equation}
\begin{split}
\mathcal{L}_{\text{m}} = &\mathcal{L}_{\text{s}} + \alpha \sum_{t=0}^{T-1} (\mathcal{L}_{\text{p}}^{t} + \mathcal{L}_{\text{h}}^{t}) + \lambda \|\theta\|^2_2
\end{split}
\end{equation}
\paragraph{Preparing the data} At training time, SPINN requires both a transition sequence $\mathbf t$  and a token sequence $\mathbf x$ as its inputs for each sentence. The token sequence is simply the words in the sentence in order. $\mathbf t$ can be obtained from any constituency parse for the sentence by first converting that parse into an unlabeled binary parse, then linearizing it (with the usual in-order traversal), then taking each word token as a \shift~transition and each `)' as a \reduce~transition, as here:

@@

A na\"ive implementation of SPINN would require simulating a stack of size $N$ for each input sentence. In order to update the model weights via backpropagation, we would also need to maintain all intermediate stack representations for later chain rule calculations. This implies a per-example space requirement of $N \times T \times D$, which is prohibitively large for significant batch sizes $M$ or sentence lengths $N$. (Recall that $T = 2N - 1$ is the number of transitions in the shift-reduce parsing sequence, and $D$ is the dimension of the vectors on the stack.) Such a naive implementation would also require duplicating a largely unchanged stack at each timestep (as \shift~and \reduce~operations only affect the top of the stack). \todo{[JG] Shorten this a tiny bit.}

@@

In its bare form, before the addition of the tracking LSTM, SPINN (in particular the SPINN-NP-NT variant, for \textit{no parsing, no tracking}) is precisely equivalent to a conventional tree-structured neural network model in the function that it computes, and it therefore also has the same learning dynamics. In both, the representation of each sentence consists of the representation of the words, combined recursively using a TreeRNN composition function (in our case, the TreeLSTM function). The SPINN, however, is dramatically faster, and supports both integrated parsing and a novel approach to context information.

@@

\begin{table*}[t]
  \centering\small
  \begin{tabular}{lcccc} 
    \toprule
Model                   & Params.    & Trans. acc.  &   Train  &   Test \\
\midrule
\multicolumn{5}{c}{Previous non-NN results}\\
\midrule
Lexicalized classifier \citep{snli:emnlp2015}
                        & --                & --                    &   99.7   &   78.2      \\
\midrule
\multicolumn{5}{c}{Previous sentence encoder-based NN results}\\
\midrule
100D LSTM encoders \citep{snli:emnlp2015}
                        & 221k               & --               &   84.8   &   77.6      \\
1024D pretrained GRU encoders \citep{DBLP:journals/corr/VendrovKFU15}
                        & 15m                & --              &   98.8   &   81.4       \\
300D Tree-based CNN encoders \citep{mou2015recognizing}
                        & 3.5m                & --             &   83.4   &   82.1       \\
\midrule
\multicolumn{5}{c}{Our results}\\
\midrule
\textbf{300D LSTM RNN encoders}          & 3.0m                  & --                &   83.9      &   80.6       \\
\textbf{300D SPINN-NP-NT encoders}
                        & 3.4m                  & --                &   84.4      &   80.9       \\
\textbf{300D SPINN-NP encoders}
                        & 3.7m                  & --                &   89.2      &   \underline{83.2}       \\
\result{\textbf{300D SPINN encoders} }
                        & ??                  & 92?           &   ?    &   82?      \\          
% \result{300D SPINN-NP, sequence-based attn.   }     
%                         & ?                  & --                &   ?      &   86.3dv         \\
% \result{300D SPINN-NP, tree-based attn. }           
%                         & ?                  & --                &   ?      &   \textbf{?}\\
\midrule
\multicolumn{5}{c}{Other previous NN results}\\
\midrule
100D LSTM w/ word-by-word attention \citep{rocktaschel2015reasoning}
                        & 252k               & --              &   85.3   &   83.5       \\
300D mLSTM word-by-word attention model \citep{DBLP:journals/corr/WangJ15b}
                        & 1.9m               & --             &   92.0   &   86.1      \\
300D LSTMN with deep attention fusion \citep{cheng2016long}
                        & 1.4m               & --                &   92.3   &   \textbf{89.1}      \\
    \bottomrule
  \end{tabular}
\caption{\protect\label{tab:results}Results on SNLI 3-way inference classification. Params. is the approximate number of trained parameters (excluding word embeddings for models where they are trained). Trans. acc. is the model's accuracy in predicting parsing transitions. Train and test are SNLI classification accuracy.} 
\end{table*}

@@

We evaluate SPINN on the task of natural language inference \citep[NLI, also known as recognizing textual entailment, or RTE][]{dagan2006pascal,MacCartney09}. NLI is a sentence pair classification task, in which a model reads two sentences (a premise and a hypothesis), and outputs a judgment of {\it entailment}, {\it contradiction}, or {\it neutral}, reflecting the relationship between the meanings of the two sentences, as in this example from the Stanford NLI corpus \citep[SNLI][]{snli:emnlp2015}, which we use for training and evaluation:

@@

Even though NLI is framed as a relatively simple three-way classification task, it is nonetheless an effective way of evaluating the ability of some model to extract broadly informative representations of sentence meaning. In order for a model to perform reliably well on NLI, it must be able to represent and reason with the core phenomena of natural language semantics, including quantification, coreference, scope, and several types of ambiguity.

@@

We evaluate four models on the SNLI task. Each uses 300D hidden states:
\begin{itemize}
\item A baseline LSTM model (similar to that of \citet{snli:emnlp2015}) that uses the same classifier architecture as our models, but encodes sentences using a single layer LSTM-RNN sequence model.
\item The minimal SPINN-NP-NT (\textit{no parsing, no tracking}), which uses transitions from an external parser and has no tracking LSTM to inform semantic composition. This model is equivalent to a TreeLSTM \citep{tai2015improved}.
\item SPINN-NP, which adds the tracking LSTM to guide composition, making it a hybrid tree-sequence model.
\item The full SPINN, which makes its own parsing decisions, and does not depend on an external parser at test time.
\end{itemize}

We compare our models against a range of baselines, including the strongest published non-neural network-based result from \citep{snli:emnlp2015}, a set of previous neural network models built around sentence encoders, and a set of models built around soft attention architectures. These attention-based models are by far the strongest published on SNLI to date, but these are not necessarily comparable to ours: SNLI was originally presented as an evaluation and development task for sentence encoding models, on the premise that sentence encoding models that do well on SNLI capture sentence meaning in a general sense, and are well suited to use on a range of NLP tasks. The attentional models are specifically adapted to the sentence-pair structure of SNLI examples and cannot act as such general sentence encoders. We focus on models which do support general sentence encoding in our evaluation.

@@

We find that the bare SPINN-NP-NT model performs little better than the RNN baseline, but that SPINN-NP with the added tracking LSTM reaches state-of-the-art accuracy among sentence encoding based models. The success of SPINN-NP, which is a hybrid tree-sequence model, suggests that the tree- and sequence-based encoding methods are complementary. The full SPINN model with its relatively weak internal parser performs slightly less well, but nonetheless exceeds the performance of the LSTM baseline. None of our models reached the performance of the strongest attention-based models, which were explicitly engineered for the sentence pair-based structure of the SNLI task. \todo{[SB] Add context here.}

Both SPINN-NP and the full SPINN outperform all previous sentence encoding models by a significant margin. Most notably our modes outperform the TBCNN of \citet{mou2015recognizing}, which also uses tree-structured composition for local feature extraction, but uses simpler pooling techniques to build sentence features in the interest of efficiency. Our results show that a model that uses tree-structured composition fully (SPINN) outperforms one which uses it only partially (TBCNN), which in turn outperforms one which does not use it at all (LSTM RNN).

@@

We implemented the feedforward of SPINN-NP-NT (which is equivalent to a standard TreeRNN) in C++/CUDA in order to compare to the performance of a standard CPU-based TreeRNN solution from \citet{irsoy2014deep}. Such a standard TreeRNN implementation operates on a single example at a time and is thus not feasible to run on a GPU.\footnote{We chose to reimplement and evaluate only the thin-stack feedforward (not training/backpropagation) in C++/CUDA, as this is the more critical performance metric for real-world applications.}

Figure~\ref{fig:speed} compares the runtime performance of the two models computing representations on data from the Stanford Sentiment Treebank \todo{cite}. Each model is restricted to run on sentences of 30 tokens or fewer (for SPINN, this means we iterate over a sequence of 59 transitions). We fix the model dimensionality and word embedding dimensionality at 300 for all tests.

@@

\todo{[JG] Note that the full SPINN (i.e., Model 1) is in the same complexity class as SPINN-NP (what we're evaluating here) with very similar constants.}
